Multi-fluids for Representing Sub-filter-scale Convection

### 2020-04-11 submit2/manuscript.tex
First complete draft with Hilary's revisions and responses to reviewers.

This paper presents \change[HW]{multi}{two}-fluid Boussinesq equations for representing
under-resolved dry convection \add[HW]{with sinking and $w=0$ air in fluid 0 and rising air in fluid 1}. We present a \remove[HW]{new} model for entrainment
and detrainment based on divergence which leads to excellent representation
of the convective area fraction. Previous multi-fluid modelling of
convection has assumed that \change[HW]{all}{both} fluids share the same pressure. \change[HW]{We show that this is}{This is shown to be} a bad assumption and \remove[HW]{we propose and validate} a model
for the pressure difference between the fluids based on divergence \add[HW]{is presented}.

Two vertical slice test cases are \change[HW]{used}{developed} to \add[HW]{tune parameters and to} \change[HW]{validate}{evaluate} the \change[HW]{multi}{use of two}-fluid
\change[HW]{model}{equations to represent dry, two-dimensional, sub-grid-scale convection}; a buoyant rising bubble and radiative convective equilibrium. \add[HW]{These}
are \add[HW]{first} simulated at high resolution and conditionally averaged based
on the sign of the vertical velocity \add[HW]{and next simulated with the two-fluid model in one vertical column}. The \change[HW]{multi}{two}-fluid model reproduces
the mean properties of the rising and falling fluids in \change[HW]{a single}{one} column. 
\end{abstract}

@@

\add[HW]{
\protect\citeA{LR01} introduced the ``assumed-distribution higher-order closure'' (ADHOC) parameterisation of the boundary layer and convection which solves prognostic equations for higher-order moments of the sub-grid-scale flow, such as the triple correlation of the vertical velocity anomalies, and diagnoses updraft and downdraft means from the higher-order moments. Thus all necessary statistics can be calculated from the grid-box-mean values that are predicted by the dynamical core. Net mass transport by convection is included because it is part of the grid-box-mean. ADHOC was extended by \protect\citeA{GLC02} to use bi-Gaussian probability distribution functions (PDFs) rather than top-hat distributions to represent sub-grid-scale variability. This formed the basis of the ``Cloud Layers Unified By Binormals'' (CLUBB) parameterization \protect\cite<>{LSW+12} which was first used at grey zone resolutions.}

\citeA{TWV+18} proposed multi-fluid modelling of convection in order
to treat non-equilibrium and net mass transport \remove[HW]{correctly and} consistently.
\change{with}{This involves using} the same numerics and physics to \change[HW]{control thermals}{prognose properties of separate parts of the fluid, for example updrafts}, the stable environment and downdrafts \remove[HW]{dynamics}. \citeA{Yano14} derived similar
equations \remove[HW]{for multiple segments} and showed how this is a generalisation
of the mass-flux formulation. The \citeA{TKP+18} ``extended eddy
diffusivity mass flux'' or extended EDMF scheme also uses similar
equations but \citeA{TKP+18}
\change[HW]{didn't propose a suitable numerical method for solving the equations.}
{only solve the equations in a single column without coupling to the continuity equation of a dynamical core.}
\change[HW]{Multi-fluids promises big improvements
in the representation of sub-grid-scale convection and unification
with the representation of turbulence, particularly at the grey zone.
However, there is a long way to go, particularly as}
{The multi-fluids approach has an advantage over ADHOC and CLUBB in solving prognostic equations for first-order moments (means of eg. updraft and environment) directly rather than solving prognostic equations for higher-order moments of the whole fluid in order to diagnose means of updrafts and the environment. Solving directly for sub-fluid means should be less sensitive to closure assumptions. However} multi-fluids require
a dedicated model rather than being a stand alone parameterisation
for a dynamical core. \citeA{TEB19} and \citeA{WM19} presented solutions
of the multi-fluid equations assuming two fluids that share the same
pressure. \remove[HW]{(or that pressure differences are predominantly due to drag
between the fluids)}. \citeA{TEB19} showed that these equations \remove[HW]{(without
drag)} are in fact unstable and can be stabilised by diffusion of vertical
velocity whereas \citeA{WM19} stabilised the equations with diffusion
and drag between fluids. Neither of these stabilisation techniques
are acceptable;
\add[HW]{sufficient diffusion to stabilise the multi-fluid equations may smooth out real features of the flow and the diffusion between fluids and drag used by \protect\citeA{WM19} meant that both fluids tended to move together rather than through each other.}
In this paper we show that there are significant
and sustained pressure differences between the fluids (section \ref{sec:results}) \add[HW]{and that the use of separate pressures stabilises the equations}.

This paper presents the multi-fluid Boussinesq equations for representing
sub-grid-scale convection including a parameterisation of the pressure
difference between fluids (section \ref{subsec:fluidPressure}) and
a new model of entrainment and detrainment (section \ref{subsec:Sij}).
The numerical method for solving the equations is described in section
\ref{sec:numerics}. We compare results of resolved dry convection
in a single-fluid two dimensional Boussinesq model with a two-fluid,
\change[HW]{single}{one} column model in section \ref{sec:results} using two vertical slice test cases.

@@

@@

\begin{table}
\begin{tabular}{c>{\raggedright}p{0.85\textwidth}}
$\sigma_{i}$ & Volume fraction of fluid $i$ so that $\sum_{i}\sigma_{i}=1$\tabularnewline
$\mathbf{u}$ & 
Velocity of the single-fluid equations  ($\text{m}\text{s}^{-1}$) and velocity averaged over all fluids of the multi-fluid equations
\tabularnewline
$\mathbf{u}_{i}$ & Velocity of fluid $i$ ($\text{m}\text{s}^{-1}$)\tabularnewline
$w_{i}$ & Vertical component of $\mathbf{u}_{i}$ ($\text{m}\text{s}^{-1}$)\tabularnewline
$S_{ij}$ & Transfer from fluid $i$ to $j$ ($\text{s}^{-1}$)
\tabularnewline
$P$ & Non-hydrostatic pressure of the single-fluid equations $=p^{\prime}/\rho_{r}$ ($\text{m}^{2}\text{s}^{-2}$) and pressure averaged over all fluids of the multi-fluid equations
\tabularnewline
$P_{i}$ & Non-hydrostatic pressure of fluid $i$ 
\tabularnewline
$p_{i}$ & $P_i - P$
\tabularnewline
$\psi_{ij}^{T}$ & Value of variable $\psi$ transferred from fluid $i$ to $j$\tabularnewline
$b$ & Buoyancy of single-fluid equations defined as -$g\rho^{\prime}/\text{\ensuremath{\rho}}_{r}$
where $\rho^{\prime}$ are departures in density from a horizontally
uniform reference, $\rho_{r}$ ($\text{m}\text{s}^{-2}$) and buoyancy averaged over all fluids of the multi-fluid equations
\tabularnewline
$b_{i}$ & Buoyancy of fluid $i$ ($\text{m}\text{s}^{-2}$)\tabularnewline
$D_{i}\big/Dt$ & Total derivative with respect to fluid $i$ $=\partial/\partial t+\mathbf{u}_{i}\cdot\nabla$\tabularnewline
$\mathbf{D}_{ij}$ & Drag on fluid $i$ from fluid $j$ ($\text{m}\text{s}^{-2}$)\tabularnewline
$\gamma$ & Coefficient for setting the pressure local for each fluid ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
$C_{D}$ & Drag coefficient $=0$ or 1\tabularnewline
$r_{c}$ & Plume radius used for defining the drag between fluids (m)\tabularnewline
$\alpha$ & Diffusivity of buoyancy ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
$\nu$ & Viscosity ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
\end{tabular}
\caption{Definitions of variables for the single- and multi-fluid Boussinesq equations.\label{tab:defns}}
\end{table}

The single-fluid Boussinesq equations will be conditionally averaged following \citeA{TWV+18} to give the multi-fluid Boussinesq equations. Lagrangian labels for each fluid $i$ are defined as $I_i=1$ in fluid $i$ and zero elsewhere. In this paper, $I_0$ labels sinking and $w=0$ fluid and $I_1$ labels rising fluid. $I_i$ satisfies:
\begin{equation}
\frac{DI_i}{Dt}=0.
\label{eq:LagrangianLabel}
\end{equation}
Applying conditional volume average to $I_i$ we define the volume fraction of fluid $i$ and the velocity and buoyancy of fluid $i$:
\begin{eqnarray}
\sigma_i &=& \overline{I_i} \\
\sigma_i \mathbf{u}_i &=& \overline{I_i \mathbf{u}} \\
\sigma_i b_i &=& \overline{I_i b}
\label{eq:defineFluidFields}
\end{eqnarray}
which implies
\begin{equation}
\sum_{i}\sigma_{i}  =  1.
\label{eq:sumOne}
\end{equation}
As we are considering the Boussinesq equations we assume that variations in density only affect the buoyancy terms so $\mathbf{u}_i$ and $b_i$ are defined without density weighting.

@@

Next we will consider conditionally averaging the terms of the momentum (\ref{eq:singleMom}) and buoyancy (\ref{eq:singleb}) equations. Conditionally averaging the advection terms leads to sub-filter-scale fluxes, $F_{SF}$, since the velocity and buoyancy are not uniform within each fluid at the filter scale:
\begin{eqnarray}
\overline{I_i \nabla \cdot (\mathbf{u} \mathbf{u})} &=& \nabla \cdot (\sigma_i \mathbf{u}_i  \mathbf{u}_i) + F_\text{SF}^{\mathbf{u}_i} \\
\overline{I_i \nabla (\mathbf{u} b)} &=&  \nabla \cdot (\sigma_i \mathbf{u}_i b_i) + F_\text{SF}^{b_i}
\label{eq:filterAdvection}
\end{eqnarray}
The sub-filter-scale fluxes will be absorbed into the diffusion terms for momentum and buoyancy.

\note[HW]{Dan, can we do better than this?}
Lots of terms appear when attempting to conditionally average the diffusion terms which we do not know how to parameterise. Therefore we will use simply:
\begin{eqnarray}
\overline{I_i \nu \nabla^2 \mathbf{u}} &=& \sigma_i \nu \nabla^2 \mathbf{u}_i \\
\overline{I_i \alpha \nabla^2 b} &=& \sigma_i \alpha \nabla^2b_i .
\end{eqnarray}

Unlike \citeA{TWV+18} we define separate pressures for each fluid, $P_i$. The pressures are defined to satisfy:
\begin{equation}
\overline{I_i \nabla P} = \sigma_i \nabla P_i.
\end{equation}

Combining these gives the conditionally averaged momentum and buoyancy equations:
\begin{eqnarray}
\frac{\partial\sigma_i\mathbf{u}_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} \mathbf{u}_{i})
+ \sigma_i \nabla P_{i}
& = &
\sigma_i b_{i}\mathbf{k}
+
\nu\sigma_i\nabla^{2}\mathbf{u}_{i}
+
\sum_{j\ne i}\left\{ \sigma_{j}S_{ji}\mathbf{u}_{ji}^{T}
 - \sigma_{i}\mathbf{D}_{ij}\right\}
\label{eq:momFlux}\\
\frac{\partial\sigma_i b_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} b_{i})
& = &
\alpha\sigma_i\nabla^{2}b_{i}
+
\sum_{j\ne i} \sigma_{j}S_{ji} b_{ji}^{T}.
\label{eq:bFlux}
\end{eqnarray}
These equations include mass transfers between fluids with $\mathbf{u}_{ij}^T$ being the velocity of the fluid transferred from $i$ to $j$ and $b_{ij}^T$ being the buoyancy of the fluid transferred from $i$ to $j$. A drag term, $\mathbf{D}_{ij}$ is also included to parameterise some of the missing terms because it is commonly used when conditional averaging is applied in fields such as multi-phase modelling \cite<eg>{GBB+07}. Sensitivity to $\mathbf{D}_{ij}$ will be investigated in section \ref{sec:results}.

@@

\citeA{TEB19} and \citeA{WM19} assumed that each \remove[HW]{co-located} fluid
shared the same pressure \change[HW]{leading}{field. This lead} to unstable equations that \citeA{TEB19}
stabilised using diffusion of vertical velocity and \citeA{WM19}
stabilised using mass exchanges or drag between the fluids. \citeA{TEB19}
achieved realistic results in \change[HW]{single}{one} column experiments of \add[HW]{the} convective
boundary layer but the amount of diffusion needed would spoil simulations
in the free troposphere \add[HW]{by diffusing real features as well as spurious artifacts}. The stabilisation used by \citeA{WM19} meant
that the two fluids tended to move as one which defeats the purpose
of using multi-fluid modelling.

@@

We will assume that pressure in each fluid is given by a \remove[HW]{parameterised}
perturbation, $p_{i}$, from the total pressure, $P$:
\begin{equation}
P_{i}=P+p_{i}.
\end{equation}
\add[HW]{In order to derive a suitable form of parameterisation for $p_i$ we will first consider the} \citeA{BN86} \add{model.}

@@

\citeA{BN86} proposed a parameterisation for pressure differences
in a multi-phase explosion model that has been widely used in mathematics
and engineering for simulating, \add[HW]{for example}, fluidised beds of granular material \cite<e.g.>{EHM92} and two compressible fluids \cite<e.g.>{SA99}. \add[HW]{This model is not used here but is examined in order to derive a method suitable for atmospheric convection.}
\citeA{SA99} wrote transport equations for both the volume fraction,
$\sigma_{i}$, and for the mass fraction,
$\sigma_{i}\rho_{i}$  without transfers between fluids which in our notation are:
\begin{eqnarray}
\frac{\partial\sigma_{i}}{\partial t}+\mathbf{V}\cdot\nabla\sigma_{i} & = & \mu\left(P_{i}-P_{j}\right)
\label{eq:BNcompaction}
\\
\frac{\partial\sigma_{i}\rho_{i}}{\partial t}+\nabla\cdot\left(\sigma_{i}\rho_{i}\mathbf{u}_{i}\right) & = & 0
\label{eq:rhoTransport}
\end{eqnarray}
where $\mathbf{V}$ is the velocity of the interface between fluids,
$\mu$ is the compaction viscosity which controls how quickly the
pressure of each fluid relaxes to the mean pressure and index $j$
refers to the other fluid in a two fluid system. In our Boussinesq
system $\rho_{i}$ does not appear in the continuity equation. Different
authors make different assumptions about the interface velocity but
if we assume that $\mathbf{V}=\sum_{i}\sigma_{i}\mathbf{u}_{i}$ and
$\rho_{i}$ is constant then eqns (\ref{eq:divFree}), (\ref{eq:sumOne}), (\ref{eq:BNcompaction}) and (\ref{eq:rhoTransport})
imply that $\mu p_{i}=-\sigma_{i}^{2}\nabla\cdot\mathbf{u}_{i}$.
If we assume that $\mathbf{V}=\mathbf{u}_{i}$ for each fluid then
we get $\mu p_{i}=-\sigma_{i}\nabla\cdot\mathbf{u}_{i}$.

@@

\citeA{TEB19} stabilised the multi-fluid equations by adding a diffusion
term $\nu\partial w^{2}/\partial z^{2}$ to the RHS of the $w$ equation.
If a fluid perturbation pressure is set to $-\nu\partial w/\partial z$
this is equivalent to the diffusion term in one dimension. However
adding a diffusion term will harm three dimensional simulations of
resolved convection in the free troposphere whereas adding a fluid
perturbation pressure will not. Inspired by both \citeA{BN86} and
\citeA{TEB19} we assume that the pressure in each fluid is given
by:
\begin{equation}
p_{i}=-\gamma\nabla\cdot(\sigma_i \mathbf{u}_{i})
\label{eq:Pi_div}
\end{equation}
where $\gamma$ is a \change[HW]{free coefficient}{compressibility with units $\text{m}^s \text{s}^{-1}$. 
This parameterisation of $p_i$ has $\sigma_i$ inside the divergence in order to ensure that $\sum_i p_i =0$ and $\therefore$ that $\sum_i P_i = P$.}

@@

\add[HW]{The transfers between fluids are related to entrainment, $\varepsilon$, and detrainment, $\delta$. Fluid 0 is the environment and fluid 1 is updrafts implying:}
\begin{equation}
S_{01} = w_1\varepsilon, \;\;\;\;\;\;\;
S_{10} = w_0\delta.
\end{equation}
\add[HW]{Results using two formulations of entrainment are presented in section \protect\ref{sec:results}. The first is}
\remove[HW]{We consider dry air without turbulence so use}
the dynamic entrainment described by
\citeA{HC51}, \citeA{AK67} and \citeA{DBF+13}:
\begin{equation}
S_{ij}=\begin{cases}
-\nabla\cdot\mathbf{u}_{i} & \text{if }\nabla\cdot\mathbf{u}_{i}<0\\
0 & \text{otherwise.}
\end{cases}
\end{equation}
\change[HW]{which is necessary to stop air in}{This formulation prevents} one fluid from accumulating when vertical motion ceases (at the top of a rising plume or at the bottom of descending air). \remove[HW]{We will also test a}{A} common form of lateral entrainment \cite<eg.>{DBF+13} \add[HW]{is also tested in section \protect\ref{sec:results}}:
\begin{equation}
\varepsilon=\frac{0.2}{r_{c}}
\end{equation}
where $r_{c}$ is the cloud or plume radius. 
\remove[HW]{Fractional entrainment is related to mass transfers by:}

@@

>>>>

This paper presents \change[HW]{multi}{two}-fluid Boussinesq equations for representing under-resolved dry convection
\add[HW]{with sinking and $w=0$ air in fluid 0 and rising air in fluid 1}. \change[HW]{We present a new}{A} model for entrainment
and detrainment based on divergence \remove[HW]{which} leads to excellent representation of the convective area fraction.
Previous multi-fluid modelling of convection has
\change[HW]{assumed that all fluids share the same pressure.}
{used the same pressure for both fluids.}
\change[HW]{We show that this is}{This is shown to be} a bad \change[HW]{assumption}{approximation} and \remove[HW]{we propose and validate} a model
for the pressure difference between the fluids based on divergence \add[HW]{is presented}.

Two vertical slice test cases are \change[HW]{used}{developed} to \add[HW]{tune parameters and to} \change[HW]{validate}{evaluate} the \change[HW]{multi}{two}-fluid \change[HW]{model}{equations for representing dry, two-dimensional, sub-grid-scale convection}; a buoyant rising bubble and radiative convective equilibrium. \add[HW]{These}
are \add[HW]{first} simulated at high resolution and conditionally averaged based
on the sign of the vertical velocity. \add[HW]{The test cases are next simulated with the two-fluid model which} \remove[HW]{The multi-fluid model} reproduces
the mean properties of the rising and falling fluids in \change[HW]{a single}{one} column. 
\end{abstract}

@@

\add[HW]{\protect\citeA{LR01} introduced the ``assumed-distribution higher-order closure'' (ADHOC) parameterisation of the boundary layer and convection which solves prognostic equations for higher-order moments of the sub-grid-scale flow, such as the triple correlation of the vertical velocity anomalies, and diagnoses updraft and downdraft means from the higher-order moments. Thus all necessary statistics can be calculated from the grid-box-mean values that are predicted by the dynamical core. Net mass transport by convection is included because it is part of the grid-box-mean. ADHOC was extended by \protect\citeA{GLC02}
to use bi-Gaussian probability distribution functions rather than top-hat distributions to represent sub-grid-scale variability. This formed the basis of the ``Cloud Layers Unified By Binormals'' (CLUBB) parameterization 
\protect\cite<>{LSW+12}
 which was first used at grey zone resolutions.}

\citeA{TWV+18} proposed multi-fluid modelling of convection in order
to treat non-equilibrium and net mass transport \remove[HW]{correctly and} consistently.
\add[HW]{The multi-fluids approach has an advantage over ADHOC and CLUBB in solving prognostic equations for first-order moments directly (means of eg. updraft and environment) rather than solving prognostic equations for higher-order moments of the whole fluid in order to diagnose means of updrafts and the environment. Solving directly for sub-fluid means should be less sensitive to closure assumptions.}
\remove{with the same numerics and physics to prognose properties of separate parts of the fluid, for example updrafts, the stable environment and downdrafts dynamics}. \citeA{Yano14} derived similar
equations \remove[HW]{for multiple segments} and showed how this is a generalisation
of the mass-flux formulation. The \citeA{TKP+18} ``extended eddy
diffusivity mass flux'' or extended EDMF scheme \remove[HW]{also} uses similar
equations but \citeA{TKP+18}
\change[HW]{didn't propose a suitable numerical method for solving the equations.}
{only solve the equations in a single column without coupling to the continuity equation of a dynamical core.}
\change[HW]{Multi-fluids promises big improvements
in the representation of sub-grid-scale convection and unification
with the representation of turbulence, particularly at the grey zone.
However, there is a long way to go, particularly as multi-fluids}
{The disadvantage of the multi-fluids approach is that it} requires
a dedicated model rather than being a stand alone parameterisation
for a dynamical core. \citeA{TEB19} and \citeA{WM19} presented solutions
of \change[HW]{the multi}{two}-fluid equations
\change[HW]{assuming two fluids that share}{using} the same pressure
\add[HW]{for both fluids}.
\remove[HW]{(or that pressure differences are predominantly due to drag
between the fluids)}. \citeA{TEB19} showed that these equations \remove[HW]{(without
drag)} are in fact unstable and can be stabilised by diffusion of vertical
velocity whereas \citeA{WM19} stabilised the equations with diffusion
and drag between fluids. Neither of these stabilisation techniques
are acceptable;
\add[HW]{sufficient diffusion to stabilise the multi-fluid equations may smooth out real features of the flow and the diffusion between fluids and drag used by \protect\citeA{WM19} meant that both fluids tended to move together rather than through each other.}
In this paper we show that there are significant
and sustained pressure differences between the fluids (section \ref{sec:results}) \add[HW]{and that the use of separate pressures stabilises the equations}.

This paper presents the multi-fluid Boussinesq equations for representing
sub-grid-scale convection including a parameterisation of the pressure
difference between fluids (section \ref{subsec:fluidPressure}) and
a new model of entrainment and detrainment (section \ref{subsec:Sij}).
The numerical method for solving the equations is described in section
\ref{sec:numerics}. \change[HW]{We compare results of resolved dry convection
in a single-fluid two dimensional Boussinesq model with a two-fluid,
single column model in section {\protect\ref{sec:results}} using two vertical slice test cases. }{Results of two vertical slice test cases are presented in section {\protect\ref{sec:results}}. These are first simulated at high resolution in a single fluid model and then statistics of the flow are reproduced solving the two-fluid equations in one column.}

@@

\subsection{\add[HW]{Derivation}}

@@

\begin{table}
\begin{tabular}{c>{\raggedright}p{0.85\textwidth}}
$\sigma_{i}$ & Volume fraction of fluid $i$ so that $\sum_{i}\sigma_{i}=1$
\tabularnewline
\add[HW]{$\mathbf{u}$} & 
\add[HW]{Velocity of the single-fluid equations  ($\text{m}\text{s}^{-1}$) and velocity averaged over all fluids of the multi-fluid equations}
\tabularnewline
$\mathbf{u}_{i}$ & Velocity of fluid $i$ ($\text{m}\text{s}^{-1}$)
\tabularnewline
$w_{i}$ & Vertical component of $\mathbf{u}_{i}$ ($\text{m}\text{s}^{-1}$)\tabularnewline
$S_{ij}$ & Transfer from fluid $i$ to $j$ ($\text{s}^{-1}$)
\tabularnewline
\add[HW]{$P$} & \add[HW]{Non-hydrostatic pressure of the single-fluid equations $=p^{\prime}/\rho_{r}$ ($\text{m}^{2}\text{s}^{-2}$) and pressure averaged over all fluids of the multi-fluid equations}
\tabularnewline
$P_i$ & Non-hydrostatic pressure of fluid $i$
\tabularnewline
\add[HW]{$p_i$} & \add[HW]{$P_i - P$}
\tabularnewline
$\psi_{ij}^{T}$ & Value of variable $\psi$ transferred from fluid $i$ to $j$
\tabularnewline
$b$ & Buoyancy of single-fluid equations defined as -$g\rho^{\prime}/\text{\ensuremath{\rho}}_{r}$
where $\rho^{\prime}$ are departures in density from a horizontally
uniform reference, $\rho_{r}$ ($\text{m}\text{s}^{-2}$) \add[HW]{and buoyancy averaged over all fluids of the multi-fluid equations}
\tabularnewline
$b_{i}$ & Buoyancy of fluid $i$ ($\text{m}\text{s}^{-2}$)
\tabularnewline
$D_{i}\big/Dt$ & Total derivative with respect to fluid $i$ $=\partial/\partial t+\mathbf{u}_{i}\cdot\nabla$
\tabularnewline
$\mathbf{D}_{ij}$ & Drag on fluid $i$ from fluid $j$ ($\text{m}\text{s}^{-2}$)
\tabularnewline
$\gamma$ & \change[HW]{Coefficient for setting the pressure local for each fluid}{Compressibility for calculating $p_i$} ($\text{m}^{2}\text{s}^{-1}$)
\tabularnewline
$C_{D}$ & Drag coefficient $=0$ or 1
\tabularnewline
$r_{c}$ & Plume radius used for defining the drag between fluids (m)\tabularnewline
$\alpha$ & Diffusivity of buoyancy ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
$\nu$ & Viscosity ($\text{m}^{2}\text{s}^{-1}$)
\tabularnewline
\end{tabular}
\caption{Definitions of variables for the single- and multi-fluid Boussinesq equations.\label{tab:defns}}
\end{table}

The single-fluid Boussinesq equations will be conditionally averaged following \citeA{TWV+18} to give the multi-fluid Boussinesq equations. Lagrangian labels for each fluid $i$ are defined as $I_i=1$ in fluid $i$ and zero elsewhere. 
For the derivation of the conditionally averaged equations we will not consider relabelling of fluid \add[HW]{and so} $I_i$ satisfies:
\begin{equation}
\frac{DI_i}{Dt}=0.
\label{eq:LagrangianLabel}
\end{equation}
Applying conditional volume average to $I_i$ we define the volume fraction of fluid $i$ and the velocity and buoyancy of fluid $i$:
\begin{eqnarray}
\sigma_i &=& \overline{I_i} \\
\sigma_i \mathbf{u}_i &=& \overline{I_i \mathbf{u}} \\
\sigma_i b_i &=& \overline{I_i b}
\label{eq:defineFluidFields}
\end{eqnarray}
which implies
\begin{equation}
\sum_{i}\sigma_{i}  =  1.
\label{eq:sumOne}
\end{equation}
As we are considering the Boussinesq equations we assume that variations in density only affect the buoyancy terms so $\mathbf{u}_i$ and $b_i$ are defined without density weighting.

@@

Next we will consider conditionally averaging the terms of the momentum (\ref{eq:singleMom}) and buoyancy (\ref{eq:singleb}) equations. Conditionally averaging the advection terms leads to sub-filter-scale fluxes, $F_{SF}$, since the velocity and buoyancy are not uniform within each fluid at the filter scale:
\begin{eqnarray}
\overline{I_i \nabla \cdot (\mathbf{u} \mathbf{u})} &=& \nabla \cdot (\sigma_i \mathbf{u}_i  \mathbf{u}_i) + F_\text{SF}^{\mathbf{u}_i} \\
\overline{I_i \nabla (\mathbf{u} b)} &=&  \nabla \cdot (\sigma_i \mathbf{u}_i b_i) + F_\text{SF}^{b_i}
\label{eq:filterAdvection}
\end{eqnarray}
The sub-filter-scale fluxes will be absorbed into the diffusion terms for momentum and buoyancy.
%
\note[HW]{Dan, can we do better than this?}
Lots of terms appear when attempting to conditionally average the diffusion terms which we do not know how to parameterise. Therefore we will use simply:
\begin{eqnarray}
\overline{I_i \nu \nabla^2 \mathbf{u}} &=& \sigma_i \nu \nabla^2 \mathbf{u}_i \\
\overline{I_i \alpha \nabla^2 b} &=& \sigma_i \alpha \nabla^2b_i .
\end{eqnarray} 
%
Unlike \citeA{TWV+18} we define separate pressures for each fluid, $P_i$. The pressures are defined to satisfy:
\begin{equation}
\overline{I_i \nabla P} = \sigma_i \nabla P_i.
\end{equation}
%
Combining these gives the conditionally averaged momentum and buoyancy equations:
\begin{eqnarray}
\frac{\partial\sigma_i\mathbf{u}_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} \mathbf{u}_{i})
+ \sigma_i \nabla P_{i}
& = &
\sigma_i b_{i}\mathbf{k}
+
\nu\sigma_i\nabla^{2}\mathbf{u}_{i}
+
\sum_{j\ne i}\left\{ \sigma_{j}S_{ji}\mathbf{u}_{ji}^{T}
 - \sigma_{i}\mathbf{D}_{ij}\right\}
\label{eq:momFlux}\\
\frac{\partial\sigma_i b_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} b_{i})
& = &
\alpha\sigma_i\nabla^{2}b_{i}
+
\sum_{j\ne i} \sigma_{j}S_{ji} b_{ji}^{T}.
\label{eq:bFlux}
\end{eqnarray}
These equations include mass transfers between fluids with $\mathbf{u}_{ij}^T$ being the velocity of the fluid transferred from $i$ to $j$ and $b_{ij}^T$ being the buoyancy of the fluid transferred from $i$ to $j$. A drag term, $\mathbf{D}_{ij}$ is also included to parameterise some of the missing terms because it is commonly used when conditional averaging is applied in fields such as multi-phase modelling \cite<eg>{GBB+07}. Sensitivity to $\mathbf{D}_{ij}$ will be investigated in section \ref{sec:results}.

@@

\citeA{TEB19} and \citeA{WM19}
\change[HW]{assumed that each co-located fluid shared the same pressure leading.}
{used the same pressure for both fluids.}
\add[HW]{This led} to unstable equations that \citeA{TEB19}
stabilised using diffusion of vertical velocity and \citeA{WM19}
stabilised using mass exchanges or drag between the fluids. \citeA{TEB19}
achieved realistic results in \change[HW]{single}{one} column experiments of \add[HW]{the} convective
boundary layer but the amount of diffusion needed would spoil simulations
in the free troposphere \add[HW]{by diffusing real features as well as spurious artifacts}. The stabilisation used by \citeA{WM19} meant
that the two fluids tended to move as one which defeats the purpose
of using multi-fluid modelling.

@@

We will assume that pressure in each fluid is given by a \remove[HW]{parameterised}
perturbation, $p_{i}$, from the total pressure, $P$:
\begin{equation}
P_{i}=P+p_{i}.
\end{equation}
\add[HW]{In order to derive a suitable form of parameterisation for $p_i$ we will first consider the} \citeA{BN86} \add{model of multi-phase flow.}

@@

\citeA{BN86} proposed a parameterisation for pressure differences
in a multi-phase explosion model that has been widely used in mathematics
and engineering for simulating, \add[HW]{for example}, fluidised beds of granular material \cite<e.g.>{EHM92} and two compressible fluids \cite<e.g.>{SA99}. \add[HW]{This model is not used here but is examined in order to derive a parameterisation of pressure differences between fluids suitable for atmospheric convection.}
\citeA{SA99} wrote transport equations for both the volume fraction,
$\sigma_{i}$, and for the mass fraction,
$\sigma_{i}\rho_{i}$  without transfers between fluids which in our notation are:
\begin{eqnarray}
\frac{\partial\sigma_{i}}{\partial t}+\mathbf{V}\cdot\nabla\sigma_{i} & = & \mu\left(P_{i}-P_{j}\right)
\label{eq:BNcompaction}
\\
\frac{\partial\sigma_{i}\rho_{i}}{\partial t}+\nabla\cdot\left(\sigma_{i}\rho_{i}\mathbf{u}_{i}\right) & = & 0
\label{eq:rhoTransport}
\end{eqnarray}
where $\mathbf{V}$ is the velocity of the interface between fluids,
$\mu$ is the compaction viscosity which controls how quickly the
pressure of each fluid relaxes to the mean pressure and index $j$
refers to the other fluid in a two fluid system. In our Boussinesq
system, $\rho_{i}$ does not appear in the continuity equation. Different
authors make different assumptions about the interface velocity but
if we assume that $\mathbf{V}=\sum_{i}\sigma_{i}\mathbf{u}_{i}$ and
$\rho_{i}$ is constant then eqns (\ref{eq:divFree}), (\ref{eq:sumOne}), (\ref{eq:BNcompaction}) and (\ref{eq:rhoTransport})
imply that $\mu p_{i}=-\sigma_{i}^{2}\nabla\cdot\mathbf{u}_{i}$.
If we assume that $\mathbf{V}=\mathbf{u}_{i}$ for each fluid then
we get $\mu p_{i}=-\sigma_{i}\nabla\cdot\mathbf{u}_{i}$.

@@

\citeA{TEB19} stabilised the multi-fluid equations by adding a diffusion
term $\nu\partial w^{2}/\partial z^{2}$ to the RHS of the $w$ equation.
If a fluid perturbation pressure is set to $-\nu\partial w/\partial z$
this is equivalent to the diffusion term in one dimension. However
adding a diffusion term will harm three dimensional simulations of
resolved convection in the free troposphere whereas adding a fluid
perturbation pressure will not. Inspired by both \citeA{BN86} and
\citeA{TEB19} we assume that the pressure in each fluid is given
by:
\begin{equation}
p_{i}=-\gamma\nabla\cdot(\sigma_i \mathbf{u}_{i})
\label{eq:Pi_div}
\end{equation}
where $\gamma$ is a \change[HW]{free coefficient}{compressibility with units $\text{m}^s \text{s}^{-1}$. 
This parameterisation of $p_i$ has $\sigma_i$ inside the divergence in order to ensure that $\sum_i p_i =0$ and therefore that $\sum_i P_i = P$ which makes the equations more striaghtforward to solve.}
\add[HW]{However it should be noted that this parameterisation destroys energy conservation.}

@@

\add[HW]{The transfers between fluids are related to entrainment, $\varepsilon$, and detrainment, $\delta$. Fluid 0 is the environment and fluid 1 is updrafts implying:}
\begin{equation}
S_{01} = w_1\varepsilon, \;\;\;\;\;\;\;
S_{10} = w_0\delta.
\end{equation}
\add[HW]{Results using two formulations of entrainment are presented in section {\protect\ref{sec:results}}. The first is}
\remove[HW]{We consider dry air without turbulence so use}
the dynamic entrainment described by
\citeA{HC51}, \citeA{AK67} and \citeA{DBF+13}:
\begin{equation}
S_{ij}=\begin{cases}
-\nabla\cdot\mathbf{u}_{i} & \text{if }\nabla\cdot\mathbf{u}_{i}<0\\
0 & \text{otherwise.}
\end{cases}
\label{eq:Sdiv}
\end{equation}
\change[HW]{which is necessary to stop air in}{This formulation prevents} one fluid from accumulating when vertical motion ceases (at the top of a rising plume or at the bottom of descending air). \remove[HW]{We will also test a}{A} common form of lateral entrainment \cite<eg.>{DBF+13} \add[HW]{is also tested in section {\protect\ref{sec:results}}}:
\begin{equation}
\varepsilon=\frac{0.2}{r_{c}}
\label{eq:fracEnt}
\end{equation}
where $r_{c}$ is the cloud or plume radius. 
\remove[HW]{Fractional entrainment is related to mass transfers by:}

@@

This test case is designed to have strong updrafts in narrow plumes
and weak descent elsewhere to mimic atmospheric convection but without
the complication of moisture, phase changes and three spatial dimensions.
